Lifting CDCL to Numerical Domains

### 2016-09-25 conclusion.tex
paper update

\section{Contributions}\label{contribution}
The goal of this paper is to instantiate ACDCL with 
different relational and non-relational numerical abstract 
domains and evaluate the performance of our solver with 
propositional-level solvers. Experimental evaluation on different
benchmarks shows that ACDCL is able to suitably exploit the 
expressivity of richer abstract domains as well as precision 
of CDCL architecture. This is attributed to the intelligent decision
heuristics which exploits the high-level structure of the problem 
combined with stronger deductions and clause learning mechanism aided 
by the richer abstract domains.

>>>>

\section{Conclusions}
In this paper, we instantiate ACDCL solver with different numerical domains -- 
intervals and octagons.  Experimental evaluation on different benchmarks shows that
ACDCL is able to suitably exploit the expressivity of richer
abstract domains as well as precision of CDCL architecture.
This is attributed to the intelligent decision heuristics
which exploits the high-level structure of the problem
combined with stronger deductions and clause learning
mechanism aided by the richer abstract domains.

### 2016-09-25 contributions.tex
paper update

\section{Contributions}\label{contribution}
The goal of this paper is to instantiate ACDCL with 
different relational and non-relational numerical abstract 
domains and evaluate the performance of our solver with 
propositional-level solvers. Experimental evaluation on different
benchmarks shows that ACDCL is able to suitably exploit the 
expressivity of richer abstract domains as well as precision 
of CDCL architecture. This is attributed to the intelligent decision
heuristics which exploits the high-level structure of the problem 
combined with stronger deductions and clause learning mechanism aided 
by the richer abstract domains.

>>>>

\section{Contributions}
The goal of this paper is to instantiate ACDCL with 
different relational and non-relational numerical abstract 
domains and evaluate the performance of our program analyzer 
with program verifiers that uses propositional-level solvers.

Experimental evaluation on different benchmarks shows that 
ACDCL is able to suitably exploit the expressivity of richer 
abstract domains as well as precision of CDCL architecture. 
This is attributed to the intelligent decision heuristics 
which exploits the high-level structure of the problem 
combined with stronger deductions and clause learning 
mechanism aided by the richer abstract domains.

### 2016-09-25 deductions.tex
paper update

\section{Propagation Strategy}\label{deduction}
Let $L$ be the number of constraints generated by a domain $D$
for a program with $N$ variables. For $D$ = {\em Interval}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em Octagons}. Note that the conjunction of all 
constraints in the set $L$ forms an octagon.

@@

\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
C program & SSA \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
int main() {
 unsigned x, y;
 __CPROVER_assume(x==y);
 x++;
 assert(x==y+1);
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
cond == (x == y)
guard1 == (cond19 && guard0)
x' == 1u + x
x' == 1u + y || !guard1
guard0 == TRUE
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{C Program ad its corresponding SSA}
\label{sw-ssa}
\end{figure}

\subsection{Forward Iteration Strategy}
For forward iteration on Figure~\ref{sw-ssa}, we need atleast 
one decision to deduce BOTTOM and prove safety. The deduction 
happens from right to left in forward iteration strategy.

@@

\subsection{Chaotic Iteration Strategy}  
The set of octagonal equilities corresponding to Figurei~\ref{sw-ssa}
is shown below. For chaotic, no decision is required to deduce BOTTOM.
\begin{figure}[t]
\scriptsize
\begin{tabular}{l}
\hline
Octagonal Inequalities \\
\hline
\begin{lstlisting}
x' > 1                                                                      
-x' - y < -2                                                               
-x - x' < -2                                                               
y - x' < 0                                                                 
x - x' < 0                                                                 
y > 0                                                                       
x > 0                                                                       
-x' - y < 0                                                                
x + y > 1                                                                 
y - x < 1                                                                 
x' - y < 2                                                                 
x - y < 1                                                                 
x + y > 0                                                                 
x + x' > 0                                                                 
x' - x < 2  
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Octagonal Inequalities}
\label{sw-ssa}
\end{figure}

>>>>

\section{Propagation Strategy}
Let $L$ be the number of constraints generated by a domain $D$
for a program with $N$ variables. For $D$ = {\em Interval}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em Octagons}. Note that the conjunction of all 
constraints in the set $L$ forms an octagon.

@@

\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
C program & SSA \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
int main() {
 unsigned x, y;
 __CPROVER_assume(x==y);
 x++;
 assert(x==y+1);
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
cond == (x == y)
guard1 == (cond19 && guard0)
x' == 1u + x
x' == 1u + y || !guard1
guard0 == TRUE
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{C Program ad its corresponding SSA}
\label{swssa}
\end{figure}

\subsection{Forward Iteration Strategy}
For forward iteration on Figure~\ref{swssa}, we need atleast 
one decision to deduce BOTTOM and prove safety. The deduction 
happens from right to left in forward iteration strategy.

@@

\subsection{Chaotic Iteration Strategy}  
The set of octagonal equilities corresponding to Figurei~\ref{sw-ssa}
is shown below. For chaotic, no decision is required to deduce BOTTOM.
\begin{figure}[t]
\scriptsize
\begin{tabular}{l}
\hline
Octagonal Inequalities \\
\hline
\begin{lstlisting}
x' > 1                                                                      
-x' - y < -2                                                               
-x - x' < -2                                                               
y - x' < 0                                                                 
x - x' < 0                                                                 
y > 0                                                                       
x > 0                                                                       
-x' - y < 0                                                                
x + y > 1                                                                 
y - x < 1                                                                 
x' - y < 2                                                                 
x - y < 1                                                                 
x + y > 0                                                                 
x + x' > 0                                                                 
x' - x < 2  
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Octagonal Inequalities}
\label{octagons}
\end{figure}

### 2016-09-25 domains.tex
paper update

\section{Abstract Domains}\label{domain}
Support for Interval, Octagons

>>>>

\section{Abstract Domains}
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}.

\subsection{Interval Abstract Domain}
Domain:
Operators:
Semantics of elements:

\subsection{Octagon Abstract Domain}
Domain:
Operators:
Semantics of elements:

\subsection{Conditions for Absract Domains}
An important criterion to support clause learning in 
partial assignments domain is that elements in the 
domain must have complementation property.  Most 
numerical abstract domain does not support 
complementation operator.  Similar to the interval 
domain, octagons does not have complementation 
property, but the decomposition of octagons into 
half spaces are complementable.

APRON library is specialized for the abstract interpretation 
framework which does not require complementation 
operator.  So, we implement our own template-based 
octagon domain which supports complementation operator.

The octagon in example can be written as a conjunction of:
\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\]

The complementation of the above octagon can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-09-25 paper.tex
paper update

\title{Automated Reasoning of Relational properties using \\
Abstact Conflict Driven Clause Learning}

\Omit{
\author{Rajdeep Mukherjee\inst{1} \and
Peter Schrammel\inst{2} \and
Daniel Kroening\inst{1}}

\institute{University of Oxford, UK \and University of Sussex, UK \\
\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
\email{P.Schrammel@sussex.ac.uk}}
}
\maketitle

%===============================================================================
\begin{abstract}
%===============================================================================
%
{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program 
analysis that embeds an abstract domain inside the Conflict Driven Clause 
Learning (CDCL) algorithm. From decision procedure perspective, ACDCL is a 
SAT solver for program analysis constraints and is a strict generalisation 
of propositional CDCL solvers. From Abstract Interpretation point of view, 
ACDCL can be viewed as an abstract interpreter that uses decision and 
learning to increase the precision of abstract transformers. In this paper, 
we present a technique to generalize ACDCL solvers to reason about relational 
properties of a program by embedding relational domains like octagon domain 
inside the ACDCL solver.  Experimental evidence shows significantly less 
number of decisions and learning iterations required by ACDCL solver to 
prove programs with relational properties compared to bit-blasting and 
applying propositional SAT solvers -- thereby emphasizing the need for 
abstract reasoning using ACDCL. 
%
\end{abstract}

%===============================================================================
\input{introduction}
%===============================================================================

@@

>>>>

\title{Lifting CDCL to Numerical Domains}

\author{Rajdeep Mukherjee Peter Schrammel   Daniel Kroening}

\institute{University of Oxford, UK \and University of Sussex, UK \\
\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
\email{p.schrammel@sussex.ac.uk}}
\Omit {
\author{Rajdeep Mukherjee\inst{1} \and
Peter Schrammel\inst{2} \and
Daniel Kroening\inst{1}}

\institute{University of Oxford, UK \and University of Sussex, UK \\
\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
\email{p.schrammel@sussex.ac.uk}}
}

\maketitle

%===============================================================================
\begin{abstract}
%===============================================================================
%
The conflict driven clause learning algorithm can be seen as an 
alternation between over-approximate deduction with boolean 
constraint propagation, and under-approximate abduction, with 
conflict analysis.  {\em Abstract Conflict Driven Clause Learning} 
(ACDCL) is a novel program analysis that embeds an abstract domain 
inside the Conflict Driven Clause Learning (CDCL) algorithm. From 
decision procedure perspective, ACDCL is a SAT solver for program 
analysis constraints and is a strict generalisation of propositional CDCL 
solvers. From Abstract Interpretation point of view, ACDCL can be 
viewed as an abstract interpreter that uses decision and learning to 
increase the precision of abstract transformers.  In this paper, we 
instantiate ACDCL with different relational and non-relational numerical 
abstract domains and evaluate the performance of our program analyzer
with program verifiers that uses propositional-level solvers.  Experimental 
evidence on a set of \rmcmt{N} benchmarks shows significant drop in number 
of decisions and backjumping iterations in addition to stronger 
deductions and learned clause aided by the richer abstract domains.
%
\end{abstract}
%===============================================================================
\input{introduction}
%===============================================================================

@@

%===============================================================================
\input{domains}
%===============================================================================
